---
title: "HW6"
author: "Josie Yaroch"
editor: visual
format: pdf
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(tidytext)
library(schrute)
library(ggplot2)
```

# The Office

Let's continue to work with *The Office* text from the [schrute](https://cran.r-project.org/web/packages/schrute/index.html) package.

# Sentiment Analysis

### Over Time

Calculate the sentiment score (positive - negative) for each episode, and plot this variable over time.

```{r}
# theoffice |> 
#   mutate(text_lower = str_to_lower(text))

tidytext::sentiments

theoffice |> 
mutate(season_episode = sprintf("S%02d E%02d", season, episode)) |>    unnest_tokens(word, text) |> # Tokenize
  inner_join(sentiments, by = join_by(word))  |>  # Merge with Sentiment Lexicon
  group_by(season, episode, season_episode) |> 
  summarize(pos_sent = sum(sentiment == "positive"),
            neg_sent = sum(sentiment == "negative"),
            sent_score = pos_sent - neg_sent,
            .groups = "drop") |> 
  arrange(season, episode) |> 
  mutate(season_episode = factor(season_episode, levels = unique(season_episode))) |> 
  ggplot(aes(x = season_episode, y = sent_score, group = season, color = factor(season))) +
  geom_line() +
  geom_point() +
  scale_x_discrete(labels = function(x) sub("^S(\\d+) E\\d+$", "S\\1", x)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = NULL, color = "Season", y = "Net sentiment")

##this plot is very messy - can't read labels on x-axis
```

```{r}
theoffice |> 
mutate(season_episode = sprintf("S%02d E%02d", season, episode)) |>    
  unnest_tokens(word, text) |> # Tokenize
  inner_join(sentiments, by = join_by(word))  |>  # Merge with Sentiment Lexicon
  group_by(season, episode, season_episode) |> 
  summarize(pos_sent = sum(sentiment == "positive"),
            neg_sent = sum(sentiment == "negative"),
            sent_score = pos_sent - neg_sent,
            .groups = "drop") |> 
  arrange(season, episode) |> 
  mutate(season_episode = factor(season_episode, levels = unique(season_episode))) |> 
  # Wrap the dataset so we can compute breaks inline (used copilot for help on this)
  (\(df) {
    # only want one label per season - going to put label on first episode per season
    season_starts <- df |>
      group_by(season) |>
      slice_min(episode, with_ties = FALSE) |>
      ungroup()
    ggplot(
      df,
      aes(x = season_episode, y = sent_score, group = season, color = factor(season))
    ) +
      geom_line() +
      geom_point() +
      #one label per season
      scale_x_discrete(
        breaks = season_starts$season_episode,
        labels = sprintf("S%02d", season_starts$season)
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
      labs(x = "Season", color = "Season", y = "Sentiment Score")
  })()
```

### By Rating

Create a scatterplot of the `imdb_rating` versus the sentiment score (positive - negative), where the case is again an episode. Decide whether a linear model is a good fit to this relationship (optional, if you have had modeling classes, what model looks best?). Comment on what you find.

```{r}
#saving the changes above (sentiment score, season_episode) into new object
theoffice0 <- theoffice |> 
mutate(season_episode = sprintf("S%02d E%02d", season, episode)) |>    
unnest_tokens(word, text, drop = FALSE) |> 
  inner_join(sentiments, by = join_by(word))  |>  # Merge with Sentiment Lexicon
  group_by(season, episode, season_episode) |> 
  mutate(pos_sent = sum(sentiment == "positive"),
            neg_sent = sum(sentiment == "negative"),
            sent_score = pos_sent - neg_sent,
            .groups = "drop") |> 
  arrange(season, episode) |> 
  mutate(season_episode = factor(season_episode, levels = unique(season_episode)))

theoffice0 |> 
  ggplot(aes(x = sent_score, y = imdb_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
      labs(x = "Sentiment Score", y = "IMDB Rating")



```

```{r}
plot(theoffice0$sent_score, theoffice0$imdb_rating,
     pch = 1, cex = 1.5,
     xlab = "Sentimant Score",
     ylab = "IMDB Rating",
     cex.lab = 1.5, cex.axis = 1.5)
regOffice <- lm(theoffice0$imdb_rating~theoffice0$sent_score)
abline(regOffice,lwd=2)
summary(regOffice)
anova(regOffice)
#Based on anova table, this model appears to be a good fit (p<<0.05)
```

### Sentiment Negation

Identify at least a few negations that change the sentiment values based on the text. Use this to recreate the graphs above (or do this first, then create the graphs).

```{r}


#Sentiment Reversal RandJ - 5910 
#common negation words : not, no, none, never, nothing, nobody, nowhere, neither, nor, hardly, seldom, rarely, scarcely
negation_words0 <- c("not", "no", "none", "never", "nothing", "nobody", "nowhere", "neither", "nor", "hardly", "seldom", "rarely", "scarcely")
#none, hardly, rarely, scarcely aren't in word1
negation_words <- c("not", "no", "never", "nothing", "nobody", "nowhere", "neither", "nor", "seldom")


# looking at count of word1:
theoffice0 |> 
  # filter(name == "Romeo and Juliet") |>  # only RandJ 
  select(text) |> 
  unnest_tokens(bigram, text, "ngrams", n = 2) |> # Tokenize
  # mutate(word1 = word(bigram,1))  # one option
  separate(bigram, into = c("word1", "word2"), remove = F) |>  
  inner_join(sentiments, by = c("word2" = "word")) |>   # Merge with Sentiment Lexicon 
  count(word1, sort = T) |>  
  filter(word1 %in% negation_words)  |> 
  group_by(word1) |> 
  count(word1, sort = T)



#in bigram, if word1 %in% negation_words & word2 is a sentiment (already filtered for this), then reverse the sentiment of word2
theoffice0 |> 
  # filter(name == "Romeo and Juliet") |>  # only RandJ 
  select(text) |> 
  unnest_tokens(bigram, text, "ngrams", n = 2) |> # Tokenize
  # mutate(word1 = word(bigram,1))  # one option
  separate(bigram, into = c("word1", "word2"), remove = F) |>  
  inner_join(sentiments, by = c("word2" = "word")) |>   # Merge with Sentiment Lexicon 
  count(word1, sort = T) |>  
  filter(word1 %in% negation_words) |> 
  head(10)
  
  
  

# Adjust this to match your tokenization (see note on contractions below)


theoffice0 |>
  select(text) |>
  # Explicit token arg; consider strip_punct = FALSE to preserve "isn't"
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ", remove = FALSE) |>
  # Join sentiment for the second word of the bigram
  inner_join(sentiments, by = c("word2" = "word")) |>
  # Keep only bigrams where word1 is a negation trigger
  filter(word1 %in% negation_words) |>
  # 🔧 Flip the sentiment label INSIDE mutate(), not piping into case_when()
  mutate(
    sentiment = case_when(
      sentiment == "positive" ~ "negative",
      sentiment == "negative" ~ "positive",
      TRUE ~ sentiment
    )
  ) |>
  count(sentiment, sort = TRUE) |>
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col() +
  labs(
    title = "Sentiment of words following negation words in The Office",
    x = "Flipped sentiment of word2 after negation",
    y = "Count"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("positive" = "lightblue", "negative" = "salmon"))

#this gives count of sentiments after negation






# --- Base unigram sentiment per episode (assumes `sentiments` is the 'bing' lexicon) ---
epi_base <- theoffice |>
  mutate(season_episode = sprintf("S%02d E%02d", season, episode)) |>
  unnest_tokens(word, text) |>
  inner_join(sentiments, by = join_by(word)) |>
  group_by(season, episode, season_episode) |>
  summarise(
    pos = sum(sentiment == "positive"),
    neg = sum(sentiment == "negative"),
    .groups = "drop"
  )

# --- Negated bigrams per episode (negator + sentiment-bearing word2) ---
epi_neg <- theoffice |>
  mutate(season_episode = sprintf("S%02d E%02d", season, episode)) |>
  unnest_tokens(bigram, text, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ", remove = FALSE, fill = "right") |>
  filter(!is.na(word2)) |>
  filter(word1 %in% negation_words) |>
  inner_join(sentiments, by = c("word2" = "word")) |>
  group_by(season, episode, season_episode) |>
  summarise(
    negated_pos = sum(sentiment == "positive"),
    negated_neg = sum(sentiment == "negative"),
    .groups = "drop"
  )

# --- Combine, compute adjusted score, and bring in IMDb rating ---
theoffice00 <- epi_base |>
  full_join(epi_neg, by = c("season", "episode", "season_episode")) |>
  mutate(
    across(c(pos, neg, negated_pos, negated_neg), ~ tidyr::replace_na(.x, 0L)),
    sent_score_raw = pos - neg,
    sent_score     = sent_score_raw - 2L * negated_pos + 2L * negated_neg
  ) |>
  left_join(
    theoffice |>
      select(season, episode, imdb_rating) |>
      distinct(),
    by = c("season", "episode")
  ) |>
  arrange(season, episode) |>
  mutate(season_episode = factor(season_episode, levels = unique(season_episode)))

##final plot for question 3
theoffice00 |>
  ggplot(aes(x = sent_score, y = imdb_rating)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Episode Sentiment vs IMDb Rating (negation-aware)",
    x = "Sentiment Score (negation-adjusted)",
    y = "IMDb Rating"
  )

```

The plot shows that as sentiment score increases, IMDb rating also tends to increase. This is true in this plot and the previous plot, showing that the trend is true when you consider the effect of negation words on the sentiment score and when you do not consider negation words
